How did the introduction of Stemmed Points affect Mobility and Site Occupation during the Late Pleistocene in Korea?

### 2020-12-14 analysis/paper/paper.Rmd
deleted include=FALSE

```{r korea-map, fig.cap="Korean Paleolithic sites mentioned in the text."}
source(here::here("analysis/paper/001-draw-the-map.R"))
knitr::include_graphics(here::here("analysis/figures/001-site-map.png"))

@@

```{r volume-density, fig.cap="The main plot shows the artefact volumetric density over time. The size of data points represent the total number of artefacts from each assemblage and the colour indicates the presence of stemmed points. The inset plot shows direct comparison of artefact densities for assemblages with and without stemmed points."}
source(here::here("analysis/paper/002-artifact-volumetric-density.R"))
knitr::include_graphics(here::here("analysis/figures/002-age-by-density.png"))
```

@@

```{r volume-density-retouch, fig.cap="Main plot shows relationship between artefact density and retouch frequency. The colour of the data points indicate the age of the assemblage, point size indicates assemblage size, and shape indicates presence or absence of stemmed points. Assemblages with higher proportions of retouched pieces and lower densities indicate curated technologies, and assemblages at the other end of the spectrum indicate expedient technologies [@riel2004late]. The inset plot on the bottom left shows the proportion of retouched pieces in the assemblages over time. The inset plot on the bottom right shows the proportion of retouched pieces in the assemblages with or without stemmed points."}
source(here::here("analysis/paper/003-artifact-volumetric-density-retouch-frequency.R"))
knitr::include_graphics(here::here("analysis/figures/003-retouch-by-density.png"))
```

@@

```{r artefact-composition, fig.cap="This plot shows the composition of each assemblage. We excluded artefacts related to manufacturing processes such as pebbles, hammers, flakes and debris, artefacts appeared only in a few assemblages including point, beak shaped, and awl, and unknown and unfinished pieces. The colour represents different types of tools and the assemblages are placed in chronological order."}
source(here::here("analysis/paper/004-artifact-composition.R"))
knitr::include_graphics(here::here("analysis/figures/004-artefact-types.png"))
```

@@

```{r raw-materials, fig.cap="This plot shows the composition of raw material type in each assemblage. We excluded raw materials that are included in less than 5 assemblages including crystal, basalt, iron_ore, slate, limestone, granite, gneiss, tuff, and unidentified ones. Hornfels are shown in 4 assemblages but since it occupies more than half of the Yongsandong and Yongband assemblages, we included it in the plot. The color represents different types of raw materials, and the assemblages are placed in chronological order."}
source(here::here("analysis/paper/005-raw-materials.R"))
knitr::include_graphics(here::here("analysis/figures/005-raw-materials.png"))
```

@@

```{r site-elevation, fig.cap="The main plot shows the site elevation by the age of the deposit. Colour indicates presence/absence of stemmed points. The lines display a locally weighted regression. The inset plot shows a direct relationship between the site elevation and stemmed points."}
source(here::here("analysis/paper/006-site-elevation.R"))
knitr::include_graphics(here::here("analysis/figures/006-site-elevation.png"))
```

@@

```{r climate-model, fig.cap="Mean Annual Temperature (MAT) of the Korean Paleolithic assemblages mentioned in this research. A: Site locations on the map and MAT. B: MAT distributions for each site during the period they were occupied. C: MAT from 50ka to 10ka. The grey lines indicate the MAT for each site and the black line is the mean temperature of all sites. The light grey area in the middle of the plot indicates the duration of MIS 2, separating MIS 3 and 1, and dark grey area represents the duration of LGM. D: The relationship between MAT and site elevation. The blue line is a linear regression on elevation and MAT with the grey area showing the 95% confidence interval.t."}
source(here::here("analysis/paper/007-climate-model-plots.R"))
knitr::include_graphics(here::here("analysis/figures/007-climate-model-sites-panel-plot.png"))
```

@@

Figure \@ref(fig:spd-model) shows the four theoretical null models of demographic change for the period 50ka to 10ka, exponential, uniform, linear, and logistic, generated from the radiocarbon dates. Statistical significance of each model is computed using 200 simulations and the global p-values are 0.004, 0.004, 0.009, and 0.004, respectively. In this case, the linear model is the best fit with lower value verified by AIC <!---(Table 1)--->. However, the other models have similar patterns with a minor margin at -48.6. The radiocarbon dates fit well with the overall trajectory of the null models (grey shading) with both positive and negative deviations. SPD shows a steady growth until 35ka, stays at the peak with minor deviation 32ka, and then starts declining with frequent temporary fluctuations (relatively sharp). The beginning and ending of the SPD indicates significant negative deviation with low density of radiocarbon dates.

>>>>

```{r korea-map, fig.cap="Korean Paleolithic sites mentioned in the text.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/001-draw-the-map.R"))}
quietly(source(here::here("analysis/paper/001-draw-the-map.R")))
knitr::include_graphics(here::here("analysis/figures/001-site-map.png"))

@@

```{r volume-density, fig.cap="The main plot shows the artefact volumetric density over time. The size of data points represent the total number of artefacts from each assemblage and the colour indicates the presence of stemmed points. The inset plot shows direct comparison of artefact densities for assemblages with and without stemmed points.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/002-artifact-volumetric-density.R"))}
quietly(source(here::here("analysis/paper/002-artifact-volumetric-density.R")))
knitr::include_graphics(here::here("analysis/figures/002-age-by-density.png"))
```

@@

```{r volume-density-retouch, fig.cap="Main plot shows relationship between artefact density and retouch frequency. The colour of the data points indicate the age of the assemblage, point size indicates assemblage size, and shape indicates presence or absence of stemmed points. Assemblages with higher proportions of retouched pieces and lower densities indicate curated technologies, and assemblages at the other end of the spectrum indicate expedient technologies [@riel2004late]. The inset plot on the bottom left shows the proportion of retouched pieces in the assemblages over time. The inset plot on the bottom right shows the proportion of retouched pieces in the assemblages with or without stemmed points.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/003-artifact-volumetric-density-retouch-frequency.R"))}
quietly(source(here::here("analysis/paper/003-artifact-volumetric-density-retouch-frequency.R")))
knitr::include_graphics(here::here("analysis/figures/003-retouch-by-density.png"))
```

@@

```{r artefact-composition, fig.cap="This plot shows the composition of each assemblage. We excluded artefacts related to manufacturing processes such as pebbles, hammers, flakes and debris, artefacts appeared only in a few assemblages including point, beak shaped, and awl, and unknown and unfinished pieces. The colour represents different types of tools and the assemblages are placed in chronological order.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/004-artifact-composition.R"))}
quietly(source(here::here("analysis/paper/004-artifact-composition.R")))
knitr::include_graphics(here::here("analysis/figures/004-artefact-types.png"))
```

@@

```{r raw-materials, fig.cap="This plot shows the composition of raw material type in each assemblage. We excluded raw materials that are included in less than 5 assemblages including crystal, basalt, iron_ore, slate, limestone, granite, gneiss, tuff, and unidentified ones. Hornfels are shown in 4 assemblages but since it occupies more than half of the Yongsandong and Yongband assemblages, we included it in the plot. The color represents different types of raw materials, and the assemblages are placed in chronological order.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/005-raw-materials.R"))}
quietly(source(here::here("analysis/paper/005-raw-materials.R")))
knitr::include_graphics(here::here("analysis/figures/005-raw-materials.png"))
```

@@

```{r site-elevation, fig.cap="The main plot shows the site elevation by the age of the deposit. Colour indicates presence/absence of stemmed points. The lines display a locally weighted regression. The inset plot shows a direct relationship between the site elevation and stemmed points.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/006-site-elevation.R"))}
quietly(source(here::here("analysis/paper/006-site-elevation.R")))
knitr::include_graphics(here::here("analysis/figures/006-site-elevation.png"))
```

@@

```{r climate-model, fig.cap="Mean Annual Temperature (MAT) of the Korean Paleolithic assemblages mentioned in this research. A: Site locations on the map and MAT. B: MAT distributions for each site during the period they were occupied. C: MAT from 50ka to 10ka. The grey lines indicate the MAT for each site and the black line is the mean temperature of all sites. The light grey area in the middle of the plot indicates the duration of MIS 2, separating MIS 3 and 1, and dark grey area represents the duration of LGM. D: The relationship between MAT and site elevation. The blue line is a linear regression on elevation and MAT with the grey area showing the 95% confidence interval.t.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/007-climate-model-plots.R"))}
quietly(source(here::here("analysis/paper/007-climate-model-plots.R")))
knitr::include_graphics(here::here("analysis/figures/007-climate-model-sites-panel-plot.png"))
```

@@

Figure \@ref(fig:spd-model) shows the four theoretical null models of demographic change for the period 50ka to 10ka, exponential, uniform, linear, and logistic, generated from the radiocarbon dates. Statistical significance of each model is computed using 200 simulations and the global p-values are 0.004, 0.004, 0.009, and 0.004, respectively. In this case, the linear model is the best fit with lower value verified by AIC. However, the other models have similar patterns with a minor margin at -48.6. The radiocarbon dates fit well with the overall trajectory of the null models (grey shading) with both positive and negative deviations. SPD shows a steady growth until 35ka, stays at the peak with minor deviation 32ka, and then starts declining with frequent temporary fluctuations (relatively sharp). The beginning and ending of the SPD indicates significant negative deviation with low density of radiocarbon dates.

```{r spd-model, fig.cap="Summed probability distribution of 108 radiocarbon dates. The black solid line represents actual radiocarbon ages. The grey shaded region shows the critical envelope that encompasses the 95% confidence interval for the null models. The red and blue vertical bands highlight the portions of the SPD where positive and negative deviations are detected. We applied logistic, exponential and linear to AIC along with the logistical growth model (fit) to evaluate each model [@riris2019widespread]. We excluded the uniform model because it has no fitted curve or surface to evaluate. The best fit model has the lowest AIC score. In this case, the linear model is the best fit, with the next-best fit (exponential) having a higher AIC score by a minor margin at -48.6.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/008-summed-probability-distributions.R"))}
quietly(source(here::here("analysis/paper/008-summed-probability-distributions.R")))
knitr::include_graphics(here::here("analysis/figures/008-summed-probability-distribution-models.png"))
```
